---
title: "Urban Sustainability and Bike-Sharing: A Global Perspective"
author: "Candidate Number: 43182"
date: "[Insert the date of your final version here]"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Prompt:** 1

**ChatGPT/AI disclosure statement:** ChatGPT was used to help brainstorm ideas and clarify concepts. However, all data collection, analysis, and writing were carried out independently.

## 1. Introduction 

You should begin your report with a concise and engaging introduction. At a minimum, this introduction should describe the data you are collecting, why it is important and interesting, and what it could be used for. Your introduction should also narrow the scope of the prompt in a way that is both clear and engaging to the reader. For example, if you choose prompt 3, your introduction should make clear what organisation you work for and what objective it is struggling to meet.


## 2. Primary Data Collection

Pulling information about city bikes:

```{r}
# Load required libraries
library(httr)
library(jsonlite)
library(dplyr)

# Base API URL
base_url <- "http://api.citybik.es/v2/networks"

# Initialize an empty data frame to store the results
citybike_info <- data.frame()

# List of network names you are interested in
desired_networks <- c("Nilesplit", "LinkBike")  # Add more network names here

# Fetch all networks to get a list of all available networks globally
response <- GET(base_url)

# Check if the response is in JSON format
if (http_type(response) == "application/json") {
  networks_data <- content(response, as = "parsed", simplifyDataFrame = FALSE)
  
  # Make sure the networks_data is a list and contains the 'networks' key
  if ("networks" %in% names(networks_data)) {
    networks_list <- networks_data$networks
    
    # Loop through each network and fetch its detailed information if it's in the desired list
    for (network in networks_list) {
      network_name <- network$name
      
      # Check if the network is in the list of desired networks
      if (network_name %in% desired_networks) {
        network_id <- network$id
        
        # Construct the URL for the specific network's detailed information
        network_url <- paste0(base_url, "/", network_id)
        
        # Fetch detailed information about the network
        network_response <- GET(network_url)
        network_details <- content(network_response, as = "parsed", simplifyDataFrame = TRUE)
        
        # Extract station details and add a new column for the network name
        stations <- network_details$network$stations
        stations$network_name <- network_name
        
        # Combine the new stations data with the citybike_info dataset
        citybike_info <- bind_rows(citybike_info, stations)
      
        print(citybike_info)
      }
    }
  } else {
    cat("Error: 'networks' not found in the API response.\n")
  }
} else {
  cat("Failed to retrieve data from the CityBikes API.\n")
}

```

You should explain what data you are collecting, how you are collecting it and why it can be considered your "primary" data.

You should explain whether there are any limits on the data you collected, and how you complied. You should also describe any ethical concerns or considerations in your data collection (e.g., when web scraping) and how you dealt with that in your code to ensure a responsible workflow.


## 3. Secondary Data Collection (energy company stock information)

You should provide a good explanation for why it is supplements the primary data, and how you envision it could be used to augment the primary data. When we evaluate your data collection in sections 2 and 3, we want to see that you have used a broad range of techniques from this course


## 4. Tabular data and transformations

[The text and code for this section goes here.]

## 5. Data Visualisation

[The text and code for this section goes here.]

## 6. Data output and storage 

[The text and code for this section goes here.]